Approximating the Optimal State Abstraction

### 2016-05-03 icml_workshop/abstract.tex

Abstraction is among the most fundamental tools to intelligent agents operating in complex spaces; through abstraction intelligent agents may pare down otherwise prohibitively large problems into tractably solvable ones. Unsurprisingly, a great deal of work in planning and \ac{RL} has focused on abstraction in \ac{MDP} in the form of state aggregation and temporally extended actions. However, there exists little work detailing what good abstractions ought to formally accomplish. In this work we propose three desiderata on abstractions in the context of planning and \ac{RL}. We summarize existing work on unifying state aggregation methods, propose an extension of this work to temporally extended actions and suggest how to formally ground out these three desiderata in our extended framework. \enote{Maybe also add experimental results, Dave?}

>>>>

Abstraction is among the most fundamental tools to intelligent agents operating in complex spaces; through abstraction, intelligent agents may pare down otherwise prohibitively large problems into tractably solvable ones. Unsurprisingly, a great deal of work in planning and \ac{RL} has focused on abstraction in \ac{MDP} in the form of state aggregation and temporally extended actions. However, there exists little work detailing what good abstractions ought to formally accomplish. In this work we propose three desiderata on abstractions in the context of planning and \ac{RL} to motivate future research. We summarize existing work on unifying state aggregation methods, and discuss how our recent work in approximate state abstraction relates to these three desiderata.

### 2016-05-03 icml_workshop/desiderata.tex

We propose the following three desiderata on theories of abstraction in the context of planning or learning, given a Markov Decision Process, $M_G$:
\begin{enumerate}
\item \textsc{Compressibility}: An abstraction ought to compress $M_G$ to be significantly smaller. \footnote{where smallness is understood as in \cite{littman1995complexity}, with respect to $|\mathcal{S}|$, $|\mathcal{A}|$, and the information content required to represent input-output pairs of $\mathcal{T}$ and $\mathcal{R}$.}
\item \textsc{Optimality}: An abstraction ought to induce an MDP $M_A$ such that solving $M_A$ basically solves $M_G$.
\item \textsc{Efficient-Computability}: An abstraction ought to consist of information that is efficient to compute.
\end{enumerate}

% Defense of the desiderata.
Abstraction, more generally, is aimed at reducing the dimensionality of an entity; this {\it reduction} is captured by \textsc{Compressibility}, while the {\it preservation} of critical information content of the entity is captured by \textsc{Optimality}. Relevant to planning and RL is the third criteria; that the information given be easy to compute. We take these three statements as guiding principles on abstractions for planning and \ac{RL} \dnote{Note about how to satisfy them?}.

Critically, all three desiderata are trivial to satisfy on their own. Given an \ac{MDP}, $M_G$, we can satisfy \textsc{Compressibility} by replacing $S_G$ with a state space that consists of a single (arbitrary) state and a single (arbitrary action). Clearly, such a resulting \ac{MDP} satisfies our first desiderata - we are left with an MDP that is substantially smaller than the original MDP. Now consider \textsc{Optimality}. Again, consider the following uninteresting method; generate an MDP $M_A$, such that $M_A$ is identical to $M_G$. Thus, the optimal policy for $M_A$ is exactly the optimal policy for $M_G$, consequently satisfying \textsc{Optimality}. Lastly for \textsc{Efficient-Computability}, returning an MDP $M_A$, such that $M_A$ is identical to $M_G$ is trivially efficiently computable.

The interesting (and difficult) case is satisfying {\it all three} desiderata. To the best of our knowledge, no abstraction has been formally proven to satisfy all three. In the next section, we discuss prior methods of abstraction, and their statuses relative to these desiderata. After, we propose an extension of some prior work that suggests how such results may be obtained for temporal abstraction.

>>>>

We propose the following three desiderata on abstraction in the context of planning or learning, given a Markov Decision Process, $M_G$:
\begin{enumerate}
\item \textsc{Compressibility}: An abstraction ought to compress $M_G$ to be significantly smaller. \footnote{where smallness is understood as in \cite{littman1995complexity}, with respect to $|\mathcal{S}|$, $|\mathcal{A}|$, and the information content required to represent input-output pairs of $\mathcal{T}$ and $\mathcal{R}$.}
\item \textsc{Optimality}: An abstraction ought to induce an MDP $M_A$ such that solving $M_A$ basically solves $M_G$.
\item \textsc{Efficient-Computability}: An abstraction ought to consist of information that is easier to compute than $\pi^*$ for $M_G$.
\end{enumerate}

% Defense of the desiderata.
Abstraction, more generally, is aimed at reducing the dimensionality of an entity; this {\it reduction} is captured by \textsc{Compressibility}, while the {\it preservation} of critical information content of the entity is captured by \textsc{Optimality}. The third criteria is relevant to planning and RL; that the information relevant to the abstraction be easy to compute. We take these three statements as guiding principles on abstractions for planning and \ac{RL}. \dnote{Note about how to satisfy them? What does it look like for one to be satisfied?}

All three desiderata are trivial to satisfy on their own. Given an \ac{MDP}, $M_G$, we can satisfy \textsc{Compressibility} by replacing $S_G$ with a state space that consists of a single (arbitrary) state and a single (arbitrary action). Clearly, such a resulting \ac{MDP} satisfies our first desiderata - we are left with an MDP that is substantially smaller than the original MDP. Now consider \textsc{Optimality}. Again, consider the following uninteresting method; generate an MDP $M_A$, such that $M_A$ is identical to $M_G$. The optimal policy for $M_A$ is exactly the optimal policy for $M_G$, consequently satisfying \textsc{Optimality}. Lastly for \textsc{Efficient-Computability}, returning an MDP $M_A$, such that $M_A$ is identical to $M_G$ is trivially efficiently computable.

The case of interest is an abstraction that satisfies {\it all three} desiderata.\footnote{Note that satisfying any {\it two} criteria is also trivial} To the best of our knowledge, no existing abstraction has been formally proven to satisfy all three desiderata (or something similar to all three).

In the next section, we discuss prior methods of abstraction, and their statuses relative to these desiderata. After, we propose an extension of some prior work that suggests how such results may be obtained for temporal abstraction.

### 2016-05-03 icml_workshop/related_work.tex

\section{Related Work}

Considerable work in planning and \ac{RL} has investigated abstraction. We group prior literature into approaches of state aggregation, temporally extended actions, or function approximation.

@@

@@

% Subsection: Abs as Temp Ext.
\subsection{Temporal Abstraction}
Previous \textit{macro-action} or \textit{options}

The use of function approximators for estimating the value function, model, or policy have seen extraordinary success in a variety of domains~\cite{mnih2015human,sutton1999policy,baird1995residual,stadie2015incentivizing,ormoneit2002kernel}. The use of such techniques achieves abstraction of a certain form - states that share some number of state variables receive similar signal from individual learning instances, thus generalizing knowledge far beyond an individual state. While some theoretical work has been done in this area~\cite{schoknecht2002optimality,thrun1993issues,parr2008analysis}, such analysis involves highly constraining assumptions (such as linearity or smoothness of the function class). Given that the goal of this work is to characterize effective abstractions by evaluating their relative satisfaction of each of our three desiderata, we stipulate that such analysis will prove illusive for these approaches.

% Subsection: Exact Abstraction Framework
\subsection{Exact Abstraction Framework}

\citet{li2006towards} developed a framework for exact state abstraction in \acp{MDP}. In particular, the authors defined five classes of state-aggregation functions, inspired by existing methods for state aggregation in \acp{MDP}. We generalize two of these five classes, $\phi_{Q^*}$ and $\phi_{\text{model}}$, to the approximate abstraction case. Our generalizations are equivalent to theirs when exact criteria are used. Additionally, when exact criteria are used our bounds indicate that no value is lost, which is one of core results of \citet{li2006towards}.

>>>>

\section{Survey of Abstraction Methods}

Considerable work in planning and \ac{RL} has investigated abstraction. We group prior literature into approaches of temporal abstraction, function approximation, and state aggregation and discuss the work's relationship to the desiderata.

% Subsection: Abs as Temp Ext.
\subsection{Temporal Abstraction}
A major agenda has been developing a formal framework around skills and temporally extended actions, including options~\cite{sutton1999between,barto2004intrinsically} and macro-actions~\cite{hauskrecht1998hierarchical}.~\cite{konidaris2007building} developed a method for building options that are transferable between similar tasks.~\cite{konidaris2014constructing} leveraged options to impose a discretized state space onto a continuous space. Notably, the resulting discretized state space, along with the given options, was shown to have the property that {\it any} plan consisting of these options would be {\it feasible}; i.e. that the sequence of operators would be executable. While this work can considerably compress a given task, no value is guaranteed to be preserved when transferring to the discretized state space. \cite{jong2008utility} demonstrate that options are not always useful in the classic four-room domain. In fact, they demonstrate that options paired with the principle of optimism under uncertainty can increase learning time. Conversely, other work has demonstrated that options can be useful in a variety of domains~\cite{sutton1999between,chentanez2004intrinsically}.

We are not aware of any work that generally characterizes what constitutes a good option, and what constitutes a bad option. Under the guise of the three desiderata, we consider this to be an important direction for future work.

The use of function approximators for estimating the value function, model, or policy have seen extraordinary success in a variety of domains with large state spaces~\cite{mnih2015human,sutton1999policy,baird1995residual,stadie2015incentivizing,ormoneit2002kernel}. The use of such techniques achieves abstraction of a certain form - states that share characteristics receive similar signal from individual learning instances, thus generalizing knowledge far beyond an individual state, also achieving a compressed representation. While some theoretical work has been done in this area~\cite{schoknecht2002optimality,thrun1993issues,parr2008analysis}, such analysis involves highly constraining assumptions (such as linearity or smoothness of the function class). Given the sorts of assumptions required, we stipulate that general analysis regarding our three desiderata will prove illusive for these approaches.

@@

\citet{li2006towards} developed a unifying framework for state abstraction approaches in \acp{MDP}. In particular, the authors defined five classes of state-aggregation functions, inspired by existing methods for state aggregation in \acp{MDP}. A state-aggregation function maps ground states of a ground MDP, $M_G$, to abstract states of an abstract MDP $M_A$. Their work provides sufficient conditions on a state aggregation function to fully maintain optimality. Consequently, these sufficient conditions satisfy \textsc{Optimality} and potentially\footnote{Degree of compressibility has not been proven, yet.} \textsc{Compressibility} are satisfied. However, the criteria proposed by \cite{li2006towards} are as difficult as solving the ground \ac{MDP} and so this approach does not satisfy \textsc{Efficient-Computability}.

@@

\enote{Add notes about how these relate to desiderata}
